---
title: "Introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Here we provide a basic introduction to the SDA4D package, an implementation of the Bayesian sparse tensor decomposition as described in brief below.  We will first explain how to generate a 4D tensor, and run a simple example, and then describe the model in some detail in order to describe the output fully.  The SDA4D package has minimal requirements; following installation no other packages should be required.

## Generating a data tensor

First, ensure you've loaded the package.
```{r setup}
#Load the library SDA4D
#library(SDA4D)
```

We provide a simple function to generate a data tensor similar to those we used in simulations, named `generate_example_data` which also returns a list of the matrices used to generate it.

```{r generatedata}
set.seed(42)
# generatedData <- generate_example_data()
# 
# dim(generatedData$dataTensor)
# 
# lapply(generateData,dim)
```

## Performing a short run of the tensor decomposition

Now, setting `stopping=FALSE` to ignore the stopping criterion we run the tensor decomposition with the function `RunSDA4D` for 100 iterations as follows:

```{r runmethod}
#res<-RunSDA4D(generatedData$dataTensor,dimn_vector = c(),maxiters = 100,stopping=FALSE)
```

In the output we have the evidence lower bound (ELBO), recorded at each iteration, the maximum number of iterations, and a number of other parameters.
```{r plotELBO,fig.width=8,fig.height=6,echo=FALSE}
res<-list(Neg_FE=exp(rnorm(100)))
# plot(log(res$Neg_FE[-1]),col='red',pch=4,xlab='Iteration',main='log(ELBO) over 100 iterations')
# lines(log(res$Neg_FE[-1]),col='red')
```

```{r plotELBOqplot,echo=FALSE,fig.width=6,fig.height=4}
library(ggplot2)
qplot(x=c(1:length(res$Neg_FE[-1])),y=res$Neg_FE[-1],geom=c("point", "line"))+xlab('Iteration')+ggtitle('log(ELBO) over 100 iterations')
```


## The model and main function output

The output of the main function is a list of a number of components:

The main RunSDA4D provides an interface to an implementation of a Bayesian sparse parallel factor analysis model for four-dimensional tensor decomposition.  

The output therefore consists of a list of the approximate posterior distributions of the main parameters, in the form of their parameters, or point estimates where these were required.  

```{r outputnames}
names(res)
```



